### 2024-05-15 project3/draft.tex
new

\subsection{Model-Based RL: From AlphaZero to Deep Hedging}
Deep reinforcement learning first attacted attentions when AlphaGo~\citep{silver2016mastering} surpasses human in playing Go. 
Soon after, AlphaZero~\citep{silver2016mastering} emerged as a model-based RL algorithm that improves over self-plays.
It quickly extended beyond Go to other games like Chess and Shogi.
In essence, the AlphaZero agent use a neural network to paramaterize the policy and the value function.
The neural network takes the current state of the game as input and outputs the probability distribution of the next move and a value of the current state.
\begin{equation}
    (\mathbf{\pi}, v) = f_{\theta}(s), \quad L(\theta) = - \mathbf{\pi}^{\intercal} \log(\mathbf{p}) + (z - v)^2  + c||\theta||^2,
\end{equation}
where $f_{\theta}$ is the neural network with parameters $\theta$, $s$ is the current state of the game, $\pi$ is the policy, $v$ is the value of the current state, and $c$ is a regularization parameter.
$\mathbf{p}$ and $z$ are the probability distribution of the next move and the outcome of the game, respectively.
They are induced by a planning algorithm, i.e., Monte Carlo Tree Search (MCTS).
Since it is infeasible to efficiently simulate trajectories in the game of Go, a planning algorithm is necessary to generate expected outcomes.
The neural network $f_{\theta}$ is trained with a stochastic gradient descent algorithm to minimize $L$.
$L$ has three components.
The last term penalizes large neural network weights.
Its first two terms measure the similarity between $f_{\theta}(s)$, the learned policy and value function, with the action and game outcome planned by the MCTS.
The neural network is trained to predict the outcome of the game and the next move before they are actually observed.
This is known as model-based RL~\citep{sutton2018reinforcement}.
Using a model allows for planning, i.e., decision-making take place using future situations that have not been experienced by the agent.

In a typical hedging problem, the transition probabilities are independent from the hedging weights, i.e., the actions.
This simplifies the planning of the neural network as complete trajectories can be simulated without the need of a planning algorithm.
In contrast to AlphaZero, deep hedging~\citep{buehler2019deep} only uses $f_{\theta}$ for planning.
More specifically, at any time $t$, the neural network $f_{\theta}$ serves as a model that predict possible future rewards before they are actually observed.
Let $(\tilde{\mathbf{s}}_0, \tilde{\Delta}_0, \tilde{\mathbf{s}}_1, \tilde{\Delta}_1, \ldots, \tilde{s}_T)$ and $\tilde{\mathcal{R}} = (R_1, \ldots, R_t)$ be a trajectory and the hedging reward of the MDP, respectively.
The neural network $f_{\theta}$ takes the previous state and hedging weight as input and outputs the hedging weight at the current state.
\begin{equation} \label{eq3:deephedging}
    \mathbf{\Delta} = f_{\theta}(\mathbf{s}_0, \ldots, \mathbf{s}_T), \quad L(\theta) = \rho(L) = \rho(-\mathcal{R}) = \rho(- \sum_{t=0}^{T} \gamma^t R_t),
\end{equation}
where $\rho(L)$ is a convex risk measure of the total present liability $L$. 
A common choice of $\rho$ is a CVaR.
In practice, the neural network $f_{\theta}$ is trained to minimize the empirical CVaR over a batch of trajectories.
It serves as a model of the hedging environment which plans according to model predictions.
~\cite{xu2020variable} propose a deep hedging algorithm for VAs that falls into this category of model-based RL.
A cost function is defined to minimize the expected shortfall of the total present liability.
~\cite{carbonneau2021deep} extends the model-based approach to hedging long-term financial derivatives and compares the performance of deep hedging strategies trained with different cost functions.
In Equation~\ref{eq3:deephedging}, the state space of the MDP contains the previous hedging weight.
A more recent paper by~\cite{imaki2021no} proposes a new network architecture that removes the previous hedging weight from the state space by introducing a no-transaction band.
It allows for a more efficient training of the neural network by removing the semi-recurrent problem structure.
Despite the success of model-based RL in hedging, most studies focus on hedging different types of contracts on classical asset models, e.g., Black-Scholes model and Heston model.
Motivated by the seminal work by~\cite{gatheral2022volatility}, there are a many subsequent works focusing on roughening the classical stochastic volatility models. 
In the rough volatility models, the latent volatility process is modelled by a stochastic process that is rougher than the sample paths of diffusion process. 
Among these rough volatility models, notable examples are rough Heston model and the rough Bergomi model. 
Model-based RL struggles in a rough volatility model due to difficulties in designing a suitable cost function.~\cite{horvath2021deep} shows that a standard implementation of deep hedging can lead to suboptimal hedging performance in a discretized rough Bergomi model. 
The CVaR-based cost function needs to be redesigned to account for jump risks, and it is still an open question.
In summary, model-based RL algorithms like deep hedging do not depend on the asset model, i.e., they are \textbf{model-independent}.
However, they are not \textbf{model-free}.
For example, deep hedging builds its own model of the hedging environment to plan the hedging strategy ahead of time using a neural network.
Designing a suitable cost function for such a network is necessary but challenging.
In contrast, a model-free approach does not require a explicit design for different asset dynamics, making it more flexible and adaptable to different market conditions and more robust to model misspecification.

@@

In a model-free reinforcement learning approach, the agent is a trial-and-error learner that interacts with the environment to learn the optimal policy by iteratively updating the policy and the value function based on the observed states, actions, and rewards.
For hedging VAs, the agent does not have access to a planning algorithm or a model of the underlying risk factors, and it must learn the optimal policy by directly interacting with the environment.
Boardly speaking, model-free RL can be categorized into two types: value-based RL and policy-based RL.
Value-based approaches learn the value function directly from the data and derive the policy from the learned value function. 
The value function is learned by training a DNN to approximate the value function, which quantifies the expected cumulative reward over time. 
The policy is then derived from the learned value function by selecting the action that maximizes the value function at each state. 
~\cite{mnih2015human} is the first paper to demonstrate the effectiveness of value-based deep reinforcement learning in training agents to play Atari games.
To update the value function, the DNN is trained to minimize the difference between the current value function and the target value function, which is updated based on the maximum expected reward from the previous iteration.
When trained to approximate the value function, the DNN benefits from the use of experience replay buffers, which is proposed by~\cite{lin1992self} to store and sample historical transitions and uniformly sampled during training to enhance the sample efficiency and stability of the learning process.
~\cite{kolm2019dynamic} uses a value-based deep reinforcement learning approach to hedge European options under the Black-Scholes framework.

@@

In the context of hedging VAs, the state space is partially observed.
The agent does not have access to the contract specifications and the underlying risk factors.
Furthermore, a rough volatility model introduces non-Markovian dynamics, i.e., the future state of the system depends on the entire history of the system.
To address the non-Markovian dynamics, an LSTM component is added to the state representation $\mathbf{s}_t$ to capture the temporal dependencies in the data.
This approach is known as recurrent PPO~\citep{ni2021recurrent}.
Let $h_t$ be the hidden state of the LSTM at time $t$.
The state representation $\mathbf{s}_t$ is fed into the LSTM at each time step to update the hidden state $h_t$.
\begin{equation} \label{eq3:stateRNN}
    \mathbf{s}_t = (S_t, F_t, G_t, \tau, \Delta_{t-1}), \quad h_t = \text{LSTM}(\mathbf{s}_t, h_{t-1})
\end{equation}

@@

The policy and reward function remain the same but with the new state representation in Equation~\ref{eq3:stateRNN}.
Instead of following the standard PPO algorithm, the agent uses the recurrent PPO algorithm to learn an optimal policy for hedging VAs.

Figure~\ref{fig3:architecturePPO} illustrates the network architecture of the standard PPO and the recurrent PPO with LSTM, and Algorithm~\ref{alg3:ppoHedging-rnn} outlines the recurrent PPO algorithm and for training an agent to hedge VAs with non-Markovian dynamics.
The recurrent PPO algorithm is similar to the standard PPO algorithm, but it includes an LSTM component to capture the temporal dependencies in the data.
However, the sequantial nature of the LSTM is not suitable for parallel computation, which slows down the training process.

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{ll} 
        \toprule
        Hyperparameter      & Value \\
        \midrule
        $|\mathcal{D}_k|$   & 64        \\
        $\epsilon$          & 0.2       \\
        Learning rate       & 0.0003    \\
        $S_0$               & 1000      \\
        $\mu$               & 0.00375   \\
        $\sigma$            & 0.0457627 \\
        $r$                 & 0.002     \\
        $T$                 & 240       \\
        \bottomrule
    \end{tabular}
    \caption{PPO Hyperparameters and GMMB Contract Specifications} 
    \label{tab3:hyperparameters}
\end{table}

>>>>

\subsection{Model-Based RL: From AlphaZero to Deep Hedging}
Deep reinforcement learning first attacted attentions when AlphaGo~\citep{silver2016mastering} surpasses human in playing Go. 
Soon after, AlphaZero~\citep{silver2016mastering} emerged as a model-based RL algorithm that improves over self-plays.
It quickly extended beyond Go to other games like Chess and Shogi.
In essence, the AlphaZero agent use a neural network to paramaterize the policy and the value function.
The neural network takes the current state of the game as input and outputs the probability distribution of the next move and a value of the current state.
\begin{equation}
    (\mathbf{\pi}, v) = f_{\theta}(s), \quad L(\theta) = - \mathbf{\pi}^{\intercal} \log(\mathbf{p}) + (z - v)^2  + c||\theta||^2,
\end{equation}
where $f_{\theta}$ is the neural network with parameters $\theta$, $s$ is the current state of the game, $\pi$ is the policy, $v$ is the value of the current state, and $c$ is a regularization parameter.
$\mathbf{p}$ and $z$ are the probability distribution of the next move and the outcome of the game, respectively.
They are induced by a planning algorithm, i.e., Monte Carlo Tree Search (MCTS).
Since it is infeasible to efficiently simulate trajectories in the game of Go, a planning algorithm is necessary to generate expected outcomes.
The neural network $f_{\theta}$ is trained with a stochastic gradient descent algorithm to minimize $L$.
$L$ has three components.
The last term penalizes large neural network weights.
Its first two terms measure the similarity between $f_{\theta}(s)$, the learned policy and value function, with the action and game outcome planned by the MCTS.
The neural network is trained to fit the planned outcomes of the game.
It stores a global solution that predicts the outcome of the game and the next move before they are actually observed.
This is known as model-based RL~\citep{sutton2018reinforcement}.
Using a model allows for planning, i.e., decision-making take place using future situations that have not been experienced by the agent.

In a typical hedging problem, the transition probabilities are independent from the hedging weights, i.e., the actions.
This simplifies the planning as complete trajectories can be simulated without the need of a planning algorithm.
In contrast to AlphaZero, deep hedging~\citep{buehler2019deep} uses the model itself to simulate the complete trajectories following the transition probabilities of the MDP.
Let $(\tilde{\mathbf{s}}_0, \tilde{\Delta}_0, \tilde{\mathbf{s}}_1, \tilde{\Delta}_1, \ldots, \tilde{s}_T)$ and $\tilde{\mathcal{R}} = (R_1, \ldots, R_t)$ be a trajectory and the hedging reward of the MDP, respectively.
The neural network $f_{\theta}$ stores a global solution, i.e., it takes the previous state and hedging weight as input and outputs the hedging weight at the current state.
\begin{equation} \label{eq3:deephedging}
    \mathbf{\Delta}_t = f_{\theta}(\mathbf{s}_t), \quad L(\theta) = \rho(L) = \rho(-\mathcal{R}) = \rho(- \sum_{t=0}^{T} \gamma^t R_t),
\end{equation}
where $\rho(L)$ is a convex risk measure of the total present liability $L$. 
The risk measure $\rho$ serves as a model of the MDP that 
A common choice of $\rho$ is a CVaR.
In practice, the neural network $f_{\theta}$ is trained to minimize the empirical CVaR over a batch of trajectories.
In other words, planning is done through $\rho$, which measures the performance of a hedging policy over a batch of complete trajectories.
The ability to know the future outcomes of the hedging policy before they are actually observed and the use of a risk measure to guide the training of the neural network are the key features of model-based RL.
For hedging VAs,~\cite{xu2020variable} propose a deep hedging algorithm that falls into this category of model-based RL.
A cost function is defined to minimize the expected shortfall of the total present liability.
~\cite{carbonneau2021deep} extends the model-based approach to hedging long-term financial derivatives and compares the performance of deep hedging strategies trained with different cost functions.
In Equation~\ref{eq3:deephedging}, the state space of the MDP contains the previous hedging weight.
A more recent paper by~\cite{imaki2021no} proposes a new network architecture that removes the previous hedging weight from the state space by introducing a no-transaction band.
It allows for a more efficient training of the neural network by removing the semi-recurrent problem structure.
Despite the success of model-based RL in hedging, most studies focus on hedging different types of contracts on classical asset models, e.g., Black-Scholes model and Heston model.
Motivated by the seminal work by~\cite{gatheral2022volatility}, there are a many subsequent works focusing on roughening the classical stochastic volatility models. 
In the rough volatility models, the latent volatility process is modelled by a stochastic process that is rougher than the sample paths of diffusion process. 
Among these rough volatility models, notable examples are rough Heston model and the rough Bergomi model. 
Model-based RL struggles in a rough volatility model due to difficulties in designing a suitable cost function.~\cite{horvath2021deep} shows that a standard implementation of deep hedging can lead to suboptimal hedging performance in a discretized rough Bergomi model. 
The CVaR-based cost function needs to be redesigned to account for jump risks, and it is still an open question.
In summary, model-based RL algorithms like deep hedging do not depend on the asset model, i.e., they are \textbf{model-independent}.
However, they are not \textbf{model-free}.
For example, deep hedging builds its own model of the hedging environment to plan the hedging strategy ahead of time using a neural network.
Designing a suitable cost function for such a network is necessary but challenging.
In contrast, a model-free approach does not require a explicit design for different asset dynamics, making it more flexible and adaptable to different market conditions and more robust to model misspecification.

@@

In a model-free reinforcement learning approach, the agent is a trial-and-error learner that interacts with the environment to learn the optimal policy by iteratively updating the policy and the value function based on the observed states, actions, and rewards.
For hedging VAs, the agent does not have access to a planning algorithm or a model of the underlying risk factors, and it must learn the optimal policy by directly interacting with the environment.
Figure~\ref{fig3:RLMethods} illustrates the critical difference between model-based and model-free RL methods.

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=\textwidth]{./project3/tikz/modelBased.pdf}
        \caption{Model-based RL (Deep Hedging)}
        \label{subfig3:mbRL}
    \end{subfigure}
    \hspace{1cm}
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=\textwidth]{./project3/tikz/modelFree.pdf}
        \caption{Model-free RL}
        \label{subfig3:mfRL}
    \end{subfigure}
    \caption{Model-based and Model-free RL Methods}
    \label{fig3:RLMethods}
\end{figure}

Boardly speaking, model-free RL can be categorized into two types: value-based RL and policy-based RL.
Value-based approaches learn the value function directly from the data and derive the policy from the learned value function. 
The value function is learned by training a DNN to approximate the value function, which quantifies the expected cumulative reward over time. 
The policy is then derived from the learned value function by selecting the action that maximizes the value function at each state. 
~\cite{mnih2015human} is the first paper to demonstrate the effectiveness of value-based deep reinforcement learning in training agents to play Atari games.
To update the value function, the DNN is trained to minimize the difference between the current value function and the target value function, which is updated based on the maximum expected reward from the previous iteration.
When trained to approximate the value function, the DNN benefits from the use of experience replay buffers, which is proposed by~\cite{lin1992self} to store and sample historical transitions and uniformly sampled during training to enhance the sample efficiency and stability of the learning process.
~\cite{kolm2019dynamic} uses a value-based deep reinforcement learning approach to hedge European options under the Black-Scholes framework.

@@

In the context of hedging VAs, the state space is partially observed.
The agent does not have access to the contract specifications and the underlying risk factors.
Furthermore, a rough volatility model introduces non-Markovian dynamics, i.e., the future state of the system depends on the entire history of the system.
To address the non-Markovian dynamics, an LSTM component is added to the state representation $\mathbf{s}_t$ to capture the temporal dependencies in the data.
This approach is known as recurrent PPO~\citep{ni2021recurrent}.
Let $h_t$ be the hidden state of the LSTM at time $t$.
The state representation $\mathbf{s}_t$ is fed into the LSTM at each time step to update the hidden state $h_t$.
\begin{equation} \label{eq3:stateRNN}
    \mathbf{s}_t = (S_t, F_t, G_t, \tau, \Delta_{t-1}), \quad h_t = \text{LSTM}(\mathbf{s}_t, h_{t-1})
\end{equation}
The difference between a feedforward neural network (FNN) and a LSTM network is illustrated in Figure~\ref{fig3:nn}.

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=\textwidth]{./project3/tikz/fnn.pdf}
        \caption{FNN}
        \label{subfig3:fnn}
    \end{subfigure}
    \hspace{1cm}
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=\textwidth]{./project3/tikz/lstm.pdf}
        \caption{LSTM}
        \label{subfig3:lstm}
    \end{subfigure}
    \caption{Neural Network Architectures}
    \label{fig3:nn}
\end{figure}

In a LSTM, the cell state $c_t$ stores a representation of historical information up to time $t$.
It is updated at each time step by a forget gate $f_t$, an input gate $i_t$, and an output gate $o_t$.

In a recurrent PPO with LSTM, the policy and reward function remain the same but with the new state representation in Equation~\ref{eq3:stateRNN}.
Instead of following the standard PPO algorithm, the agent uses the recurrent PPO algorithm to learn an optimal policy for hedging VAs.
Figure~\ref{fig3:architecturePPO} illustrates the network architecture of the standard PPO and the recurrent PPO with LSTM, and Algorithm~\ref{alg3:ppoHedging-rnn} outlines the recurrent PPO algorithm and for training an agent to hedge VAs with non-Markovian dynamics.

@@

The recurrent PPO algorithm is similar to the standard PPO algorithm, but it includes an LSTM component to capture the temporal dependencies in the data.
However, the sequantial nature of the LSTM is not suitable for parallel computation, which slows down the training process.

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{ll} 
        \toprule
        Hyperparameter      & Value \\
        \midrule
        $|\mathcal{D}_k|$   & 16        \\
        $\epsilon$          & 0.2       \\
        Learning rate       & 0.0001    \\
        $S_0$               & 1000      \\
        $\mu$               & 0.00375   \\
        $\sigma$            & 0.0457627 \\
        $r$                 & 0.002     \\
        $T$                 & 240       \\
        \bottomrule
    \end{tabular}
    \caption{PPO Hyperparameters and GMMB Contract Specifications} 
    \label{tab3:hyperparameters}
\end{table}
